\plaintitle

### 2016-09-13 drafts/labor/piecework_lit.tex
complexity should be kind of done
msb comments

\makeatletter
\makeatother
\begin{document}
\section{A Review of Piecework}
\topic{The HCI community has used the term ``piecework'' to describe
myriad instantiations of on--demand labor%
, but this reference has generally been offered in passing.}
As this paper principally traces a relationship between
the historical piecework and the contemporary crowdwork
(or on--demand labor more generally)%
, this casual familiarity with piecework may prove insufficient.
We'll more carefully discuss piecework in this section in order
to inform the subsequent sections --- and indeed, the entire argument.
Specifically, we will
\begin{inlinelist}
  \item define ``piecework'' as researchers in the topic understood it;
  \item trace the rise of piecework at a very high level%
,         identifying key figures and ideas during this time; and finally
  \item look at the fall of piecework, such as it was%
,         considering in particular
        the factors that may have led to piecework's eventual demise
\end{inlinelist}.

@@

\topic{One of the earliest definitions of piecework%
, in \citeyear{hughRaynbirdTaskWork}, also proves to be the most circumspect in its wording.}
\citeauthor{hughRaynbirdTaskWork} offers
a concise definition of piecework
--- which he variously also calls ``measure work'', ``grate work'', and ``task work'' ---
by contrasting the ``task--labourer'' with the ``day--labourer'':
``\dots the chief difference lies between the day--labourer,
who receives a certain some of money\dots~for his day's work,
and the task--labourer, whose earnings depend on the \textit{quantity} of work done [emphasis added]''
\cite{hughRaynbirdTaskWork}.
This description offers the first rudimentary definition of piecework
from which the practice will grow for more than a century;
piece work, as \citeauthor{hughRaynbirdTaskWork} offers,
``depend[s] on the quantity of work done''.

@@

\topic{It's worth acknowledging that --- as \citeauthor{hart2013rise} point out ---
``this distinction [between piece--rates and time--rates] was not completely clear--cut''
\cite{hart2013rise}.}
The ``Rowan premium system''%
, which essentially paid workers a base rate for time with the opportunity for additional pay associated with output,
was just one of several alternatives to stricter time-- and piece--rate renumeration paradigms, which
muddies the waters for us later as we attempt to categorize cases of piecework
\cite{rowan1901premium}.
Nevertheless,
this work offers us an intuition for what piecework is, if not a bright--line rule.

@@

\topic{We see this sense of independence regardless of the time, locale, and industry.}
\citeauthor{10.2307/3827491} offers us a look into the lives and culture of ``match girls''
--- young women paid by piecework to assemble matchsticks generally in the late \nth{19} century.
Of particular interest was their reputation ``\dots~for generosity, independence, and protectiveness,
but also for brashness, irregularity, low morality, and little education''
\cite{10.2307/3827491}.
\citeauthor{10.2307/27508091} documents piecework from 1850--1930 in Australia%
, finding similar assertions of the freedom compositors of newspapers experienced as piece workers:
``If a piece--work compositor who held a `frame' decided that he did not want to work on a particular day or night,
the management recognised his right to put a `substitute' or `grass' compositor in his place''
\cite{10.2307/27508091}.
From these accounts we should be able to identify
a sense of independence that
resonates across decades, industries, and locales where piecework is found.
We'll problematize this supposed advantage as we trace the history of piecework%
, but for now we can say that piecework affords
independence and some sense of locus
otherwise unknown to workers.

\topic{\citeauthor{hart2013rise} offer another series of compelling insights toward the question of the features that sprout from piecework.}
In their reflection on the features endemic to piecework in the 1930s,
which they describe as the ``heyday'' of piecework's prominence;
among them were the following:
\begin{inlinelist}
\item ``female workers who generally had less training'' had to be trained in narrower subsets of the general body of skills that conventional (male) apprentices would undertake, and
\item workers with specific slices of skills could be more appropriately matched to suitable tasks
\end{inlinelist}
\cite{hart2013rise}.

\topic{Consolidating what we've learned from these sources about piecework, we might be able to arrive at a working definition that will
suffice for our needs.}
Those needs being that the definition be
\begin{inlinelist}
\item faithful to the historical cases of piecework that we see in the scholarship; and
\item relevant and informative to potential cases of piecework today.
\end{inlinelist}
We offer the following:
that \pieceworkdefinition.

@@

\topic{Piecework's history traces back further perhaps than most would expect.}
\citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
to calculate equations in order to better--predict the trajectories of various celestial bodies
\cite{grier2013computers}.
While this approach didn't become an economic powerhouse as later examples would prove%
, Airy and others arguably found the kernel of insight that we pursue throughout this discussion:
determining the extent to which work can be decomposed, and
finding the limits of complexity of that decomposed work.
That is, Airy found that he could train youths in elementary mathematics
to complete the majority of the calculations he would otherwise have had to solve on his own%
, and that the greater body of work could ultimately be completed sooner
if he arranged his work appropriately.

\topic{Piecework took a circuitous path in its rise to the mainstream%
% from the field into the home and finally into factories%
, each time finding additional ways to leverage the advantages of piecework.}
First applied to farm work, as
\citeauthor{hughRaynbirdTaskWork} and others illustrate%
, the practice remained relatively obscure until
it was brought to the textile industry
\cite{hughRaynbirdTaskWork}.
At the turn of the \nth{20} century%
, when \citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City%
, \citeauthor{norton1900textile} was providing substantive guidance on various wage regimes%
, describing piecework comprehensively
% ``The timeworkers' wages should be verified by the timekeeper's book.''
\cite{riisOtherSideLives,norton1900textile}.
Soon after,
we saw the application of piecework systems in textile mills on the realization that
``[pieceworkers in Italy] will work as many hours as it is possible for him to stand''
\cite{clark1908cotton}.
% By the turn of the \nth{20} century%
% , \citeauthor{riisOtherSideLives} 
% the poor working and living conditions that were pervasive in
% an increasingly industrial United States
% \cite{riisOtherSideLives}.
% By the turn of the \nth{20} century,
best practices regarding the measurement and management of
piecework rates, and of workers in the engineering industry,
were beginning to take shape
\cite{burton1899commercial}.

\topic{Researchers have since struggled
to understand the mechanisms and characteristics in piecework
which fueled its rise to popularity during this time.}
\citeauthor{10.2307/23702539} argued that the first sparks of scientific management
could be found in piecework;
the approach of paying workers for each piece of output necessitated
the rigorous tracking, measurement, and training of workers
for which scientific management became famous
\cite{10.2307/23702539}.
This argument is certainly compelling;
it would seem to make the concurrent upswing of
scientific management and Fordism
through the first two--thirds of the \nth{20} century
alongside piecework not only understandable, but predictable
\cite{hart2013rise}.
\citeauthor{Brown01041990} inquired from another direction, asking
what limited the adoption of piecework in industries that otherwise gravitated toward it
(in the case studies he examined, this mostly focused on railway engineers)
\cite{Brown01041990}.

@@

\subsection{Why is piecework relevant to crowdwork?}
Using the definition of piecework that we came up with earlier, we argue that
crowdwork is fundamentally an instantiation of piecework, and
that we can more precisely anticipate the answers to the open research questions we discussed earlier.
We'll show that the dimensions of crowdwork that the broader HCI community has been studying
align with the history of piecework, and that this can greatly inform predictions about the future of crowd work.

>>>>

\makeatletter
\makeatother
\begin{document}
\section{A Review of Piecework}
\topic{The HCI community has used the term ``piecework'' to describe
myriad instantiations of on--demand labor%
, but this reference has generally been offered in passing.}
As this paper principally traces a relationship between
the historical piecework and the contemporary crowdwork \msb{consistency: crowdwork or crowd work?}
(or on--demand labor more generally)%
, this casual familiarity with piecework may prove insufficient.
We'll more carefully discuss piecework in this section in order
to inform the subsequent sections --- and indeed, the entire argument.
Specifically, we will
\begin{inlinelist}
  \item define ``piecework'' as researchers in the topic understood it;
  \item trace the rise of piecework at a very high level%
,         identifying key figures and ideas during this time; and finally
  \item look at the fall of piecework, such as it was%
,         considering in particular
        the factors that may have led to piecework's \msb{(and crowd work's?)} eventual demise
\end{inlinelist}.

@@

\topic{One of the earliest definitions of piecework%
, in \citeyear{hughRaynbirdTaskWork}, also proves to be the most circumspect in its wording.} \msb{The circumspect bit is weird, not sure how to interpret that.}
\citeauthor{hughRaynbirdTaskWork} offers
a concise definition of piecework
--- which he variously also calls ``measure work'', ``grate work'', and ``task work'' ---
by contrasting the ``task--labourer'' with the ``day--labourer'':
``\dots the chief difference lies between the day--labourer,
who receives a certain some of money\dots~for his day's work,
and the task--labourer, whose earnings depend on the \textit{quantity} of work done [emphasis added]''
\cite{hughRaynbirdTaskWork}.
This description offers the first rudimentary definition of piecework
from which the practice will grow for more than a century;
piece work, as \citeauthor{hughRaynbirdTaskWork} offers,
``depend[s] on the quantity of work done''.

@@

\topic{It's worth acknowledging that --- as \citeauthor{hart2013rise} point out ---
``this distinction [between piece--rates and time--rates] was not completely clear--cut''
\cite{hart2013rise}.}
The ``Rowan premium system''%
, which essentially paid workers a base rate for time with the opportunity for additional pay associated with output,
was just one of several alternatives to stricter time-- and piece--rate renumeration paradigms, which
muddies the waters for us later as we attempt to categorize cases of piecework
\cite{rowan1901premium}. \msb{That point runs way too fast for me --- why does it muddy the waters? Either simplify, cut, or slow it down.}
Nevertheless,
this work offers us an intuition for what piecework is, if not a bright--line rule. \msb{I'd cut that, I think it's fine...}

@@

\topic{We see this sense of independence regardless of the time, locale, and industry.}
\citeauthor{10.2307/3827491} offers us a look into the lives and culture of ``match girls''
--- young women paid by piecework to assemble matchsticks generally in the late \nth{19} century.
Of particular interest was their reputation ``\dots~for generosity, independence, and protectiveness,
but also for brashness, irregularity, low morality, and little education''
\cite{10.2307/3827491}.
\citeauthor{10.2307/27508091} documents piecework from 1850--1930 in Australia%
, finding similar assertions of the freedom compositors of newspapers experienced as piece workers:
``If a piece--work compositor who held a `frame' decided that he did not want to work on a particular day or night,
the management recognised his right to put a `substitute' or `grass' compositor in his place''
\cite{10.2307/27508091}.
From these accounts we should be able to identify
a sense of independence that
resonates across decades, industries, and locales where piecework is found.
We'll problematize this supposed advantage as we trace the history of piecework%
, but for now we can say that piecework affords
independence and some sense of locus \msb{not sure what that means...a sense of place?}
otherwise unknown to workers.

\topic{\citeauthor{hart2013rise} offer another series of compelling insights toward the question of the features that sprout from piecework.}
In their reflection on the features endemic to piecework in the 1930s,
which they describe as the ``heyday'' of piecework's prominence;
among them were the following:
\begin{inlinelist}
\item ``female workers who generally had less training'' had to be trained in narrower subsets of the general body of skills that conventional (male) apprentices would undertake, and
\item workers with specific slices of skills could be more appropriately matched to suitable tasks
\end{inlinelist}
\cite{hart2013rise}. \msb{explain that in your own words now --- that the piecework infrastructure allows lower-skilled people to be used for skilled positions, by building the expertise into the workflow}

\topic{Consolidating what we've learned from these sources about piecework, we might be able to arrive at a working definition that will
suffice for our needs.} \msb{Again, I don't think the goal for the reader should be to arrive at a definition through a narrative, they instead need to know how to think about piecework. This  paragraph could probably be cut}
Those needs being that the definition be
\begin{inlinelist}
\item faithful to the historical cases of piecework that we see in the scholarship; and
\item relevant and informative to potential cases of piecework today.
\end{inlinelist}
We offer the following:
that \pieceworkdefinition.

@@

\topic{Piecework's history traces back further perhaps than most would expect.}
\citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
to calculate equations in order to better--predict the trajectories of various celestial bodies
\cite{grier2013computers}. \msb{year?}
While this approach didn't become an economic powerhouse as later examples would prove%
, Airy and others arguably found the kernel of insight that we pursue throughout this discussion:
determining the extent to which work can be decomposed, and
finding the limits of complexity of that decomposed work.
That is, Airy found that he could train youths in elementary mathematics
to complete the majority of the calculations he would otherwise have had to solve on his own%
, and that the greater body of work could ultimately be completed sooner
if he arranged his work appropriately.

\msb{After reading this paragraph, I don't know what it's supposed to be teaching me. What I got out of it is that a bunch of people did piecework, but I don't know why or in fact why these are different than the sources we cite earlier. Can you hone it?}
\topic{Piecework took a circuitous path in its rise to the mainstream%
% from the field into the home and finally into factories%
, each time finding additional ways to leverage the advantages of piecework.} \msb{that sentence has some circular logic going on, it took a path to leveraging the advantages of itself?}
First applied to farm work, as
\citeauthor{hughRaynbirdTaskWork} and others illustrate%
, the practice remained relatively obscure until
it was brought to the textile industry
\cite{hughRaynbirdTaskWork}.
At the turn of the \nth{20} century%
, when \citeauthor{riisOtherSideLives} was documenting abhorrent working \& living conditions of pieceworkers in New York City%
, \citeauthor{norton1900textile} was providing substantive guidance on various wage regimes%
, describing piecework comprehensively \msb{here's an example of why I'm confused. there's guidance and comprehensive description, but this text doesn't tell me what actually happened with piecework.}
% ``The timeworkers' wages should be verified by the timekeeper's book.''
\cite{riisOtherSideLives,norton1900textile}.
Soon after,
we saw the application of piecework systems in textile mills on the realization that
``[pieceworkers in Italy] will work as many hours as it is possible for him to stand''
\cite{clark1908cotton}.
% By the turn of the \nth{20} century%
% , \citeauthor{riisOtherSideLives} 
% the poor working and living conditions that were pervasive in
% an increasingly industrial United States
% \cite{riisOtherSideLives}.
% By the turn of the \nth{20} century,
Best practices regarding the measurement and management of
piecework rates, and of workers in the engineering industry,
were beginning to take shape
\cite{burton1899commercial}.

\topic{Researchers have since struggled
to understand the mechanisms and characteristics in piecework
which fueled its rise to popularity during this time.}
\citeauthor{10.2307/23702539} argued that the first sparks of scientific management
could be found in piecework;
the approach of paying workers for each piece of output necessitated
the rigorous tracking, measurement, and training of workers
for which scientific management became famous
\cite{10.2307/23702539}.
This argument is certainly compelling;
it would seem to make the concurrent upswing of
scientific management and Fordism
through the first two--thirds of the \nth{20} century
alongside piecework not only understandable, but predictable
\cite{hart2013rise}.
\citeauthor{Brown01041990} inquired from another direction, asking
what limited the adoption of piecework in industries that otherwise gravitated toward it
(in the case studies he examined, this mostly focused on railway engineers)
\cite{Brown01041990}. \msb{Seems weird to leave the question hanging. Either un-questionify it, or answer it.}

@@

\subsection{Why is piecework relevant to crowdwork?}
\msb{I assume this is forthcoming?}
Using the definition of piecework that we came up with earlier, we argue that
crowdwork is fundamentally an instantiation of piecework, and
that we can more precisely anticipate the answers to the open research questions we discussed earlier.
We'll show that the dimensions of crowdwork that the broader HCI community has been studying
align with the history of piecework, and that this can greatly inform predictions about the future of crowd work.
